Predicting Success of Motivational Interviews with Recurrent Neural Network and Probabilistic Model for Patient-Provider Communication Sequences

### 2017-08-28 paper-amia-summit-2018/amia_summit_2018.tex
update abstract

\title{Predicting Success of Motivational Interviews with Recurrent Neural Network and Probabilistic Model of Patient-Provider Communication Sequences}

@@

\noindent{\bf Abstract}
\textit{The problem of analyzing temporally ordered observation sequences to make predictions related to the outcome of the processes that generated these sequences arises in many domains of healthcare informatics. In this paper, we focus on patient-provider communication sequences in the context of clinical interviews and propose two methods, based on Markov chain and Hidden Markov Model, for predicting the likelihood of eliciting a particular type of patient behavioral response based on an observed sequence of patient-provider exchanges.  Our method achieved 75.32\% and 79.89\% F-measure for natural sequences and 97.36\% and 96.99\% F-measure for alternating sequences in predicting the outcome of motivational interviews with obese adolescents using first and second order Markov Chain and Hidden Markov Model, respectively. The proposed method can be used to automatically identify the most effective communication strategies in motivational interviews, which significantly decreases the effort required to develop effective interventions to address many public health conditions.}

>>>>

\title{Predicting Success of Motivational Interviews with Recurrent Neural Network and Probabilistic Model for Patient-Provider Communication Sequences}

@@

\noindent{\bf Abstract}
\textit{The problem of analyzing temporally ordered observation sequences to make predictions related to the outcome of the processes that generated these sequences arises in many domains of healthcare informatics. In this paper, we focus on patient-provider communication sequences in the context of clinical interviews and propose Recurrent Neural Network (RNN) with two baseline methods Markov chain and Hidden Markov Model (HMM), for predicting the likelihood of eliciting a particular type of patient behavioral response based on an observed sequence of patient-provider exchanges. Our method achieved 70.03\%, 48.65\%, 88.53\% F1-score for under-sampled sequences and 78.96\%, 77.06\%, 87.16\% F1-score for over-sampled sequences in predicting the outcome of motivational interviews with obese adolescents using Markov Chain, HMM, and RNN, respectively. The proposed method can be used to automatically identify the most effective communication strategies in motivational interviews, which significantly decreases the effort required to develop effective interventions to address many public health conditions.}
